---
title: "My Lab 5"
author: "Skyler Moomey"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
```{r}
getwd()
```

## Task 2

### Binomial Simulation
```{r}
#Dr. Stewart's sampling function
## sampling function
# iter = iterations, n=sample size
# set default values
mybin=function(iter=100,n=10, p=0.5){ 
  # make a matrix to hold the samples
  #initially filled with NA's
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  #Make a vector to hold the number of successes in each trial
  succ=c()
  for( i in 1:iter){
    #Fill each column with a new sample
    sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))
    #Calculate a statistic from the sample (this case it is the sum)
    succ[i]=sum(sam.mat[,i])
  }
  #Make a table of successes
  succ.tab=table(factor(succ,levels=0:n))
  #Make a barplot of the proportions
  barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
  succ.tab/iter
}

mybin(n=10, p=0.7)
mybin(iter=200, n=10, p=0.7)
mybin(iter=500, p=0.7)
mybin(iter=1000, p=0.7)
mybin(iter=10000, p=0.7)
```


### Verification

```{r}
#verify values for table with 10000 iterations
dbinom(x=0, size=10, prob=0.7)
dbinom(x=1, size=10, prob=0.7)
dbinom(x=2, size=10, prob=0.7)
dbinom(x=3, size=10, prob=0.7)
dbinom(x=4, size=10, prob=0.7)
dbinom(x=5, size=10, prob=0.7)
dbinom(x=6, size=10, prob=0.7)
dbinom(x=7, size=10, prob=0.7)
dbinom(x=8, size=10, prob=0.7)
dbinom(x=9, size=10, prob=0.7)
dbinom(x=10, size=10, prob=0.7)
```

All of the values in the relative frequency histogram for the Binomial simulation with $10,000$ iterations are approximately their true values.

## Task 3

### Random Sample without Replacement, S=12, F=8 
```{r}
whites <- c()
for(i in 1:12) {
  whites <- c(whites, "1")
}

blacks <- c()
for(i in 1:12) {
  blacks <- c(blacks, "0")
}

population <- c(whites, blacks)

## sample
sample(population, size=5, replace=FALSE)
```

### Hypergeometric Simulation, S=12, F=8, n=5
```{r}
##hypergeometric distribution simulation function from Dr. Stewart
myhyper=function(iter=100,N=20,r=12,n=5){
  # make a matrix to hold the samples
  #initially filled with NA's
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  #Make a vector to hold the number of successes over the trials
  succ=c()
  for( i in 1:iter){
    #Fill each column with a new sample
    sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
    #Calculate a statistic from the sample (this case it is the sum)
    succ[i]=sum(sam.mat[,i])
  }
  #Make a table of successes
  succ.tab=table(factor(succ,levels=0:n))
  #Make a barplot of the proportions
  barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
  succ.tab/iter
}

myhyper()
myhyper(iter=200)
myhyper(iter=500)
myhyper(iter=1000)
myhyper(iter=10000)
```

### Verification
```{r}
#true probabilities in hypergeometric distribution
dhyper(0, 12, 8, 5)
dhyper(1, 12, 8, 5)
dhyper(2, 12, 8, 5)
dhyper(3, 12, 8, 5)
dhyper(4, 12, 8, 5)
dhyper(5, 12, 8, 5)
```


## Task 4

### mysample() Function

I believe mysample() should print relative frequency histograms for a random sample of a population of size ten of sample size n with replacement. It will create a different barplot for each iteration of the function and display them sequentially.

```{r}

# Dr. Stewart's mysample function
mysample=function(n, iter=10,time=0.5){
for( i in 1:iter){
#make a sample
s=sample(1:10,n,replace=TRUE)
# turn the sample into a factor
sf=factor(s,levels=1:10)
#make a barplot
barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)

#release the table
Sys.sleep(time)
}
}

#function call
mysample(n=1000, iter=30, time=1)
```

Each element should have an equal probability of being drawn, and we see that this is generally true.

## Task 5

### R Calculations
```{r}
choose(8, 4)
1 - ppois(3, 2)
dnbinom(7, prob=0.4, size=3)
pbinom(8, size=15, prob=0.4)
```

### Results
${8\choose4} = 70\\ P(Y > 4), Y \sim Pois(\lambda=2) \approx 0.1428765\\ P(Y=10), Y\sim NegBin(p=0.4, r = 3) \approx 0.6449725\\$
$P(Y \leq 3), Y \sim Bin(n=15, p=0.4) \approx 0.9049526$
